---
title: "Advanced R Problems - Subsetting"
author: "Michael Kozel"
date: "February 13, 2018"
output: html_document
---

# Subsetting

## Exercise 1
### Fix each of the following common data frame subsetting errors:
#### mtcars[mtcars$cyl = 4, ]
#### mtcars[-1:4, ]
#### mtcars[mtcars$cyl <= 5]
#### mtcars[mtcars$cyl == 4 | 6, ]
```{r}
mtcars[mtcars$cyl==4,] # added =
mtcars[-(1:4)] #added () and removed ,
mtcars[mtcars$cyl <= 5,] # added ,
mtcars[mtcars$cyl == 4 | 6, ] # this worked.  Maybe something has changed in R but it worked.
```

## Exercise 2
### Why does x <- 1:5; x[NA] yield five missing values? (Hint: why is it different from x[NA_real_]?)
```{r}
x <- 1:5; x[NA]
```
It returns five NAs because you are subsetting a vector of 1 through 5 by NA 

## Exercise 3
### What does upper.tri() return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?
```{r}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```
using ?upper.tri it says that it returns a matrix of logiccals the same size of a given matrix with entries TRUE in the lower of or upper triangle.

## Exercise 4
### Why does mtcars[1:20] return an error? How does it differ from the similar mtcars[1:20, ]?

```{r}
mtcars[1:20, ]
length(mtcars)
class(mtcars)
```

Error in `[.data.frame`(mtcars, 1:20) : undefined columns selected.  The reason it gives an error is because it is trying to subset a dataframe by a vector that is longer than what the mtcars dataframe is. mtcars is only 11 long, not 20. adding the , after the 20 now subsets the dataframe like a matrix.


## Exercise 5
### Implement your own function that extracts the diagonal entries from a matrix (it should behave like diag(x) where x is a matrix).

```{r}
Fun <- function(x) {
  stopifnot(is.matrix(x), nrow(x) == ncol(x))
  n <- nrow(x)
  return(x[matrix(seq_len(n), nrow = n, ncol = 2)])
}
```
https://github.com/aammd/UBCadv-r/wiki/02:-Subsetting

## Exercise 6
### What does df[is.na(df)] <- 0 do? How does it work?

It reurns an error that it is not subsettable. So it does not work

## Exercise 7
### Given a linear model, e.g., mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Extract the R squared from the model summary (summary(mod))
```{r}
mod <- lm(mpg ~ wt, data = mtcars)
summary(mod)


mod$df.residual
```

## Exercise 8
### How would you randomly permute the columns of a data frame? (This is an important technique in random forests.) Can you simultaneously permute the rows and columns in one step?
```{r}
mtcars

mtcars[setdiff(names(mtcars), "gear")] # replace gear with whatever column you wish to remove
```

## Exercise 9
### How would you select a random sample of m rows from a data frame? What if the sample had to be contiguous (i.e., with an initial row, a final row, and every row in between)?
```{r}
df <- data.frame(x = rep(1:3, each = 2), y = 6:1, z = letters[1:6])
df
df[sample(nrow(df)), ]

# you could also do
# df[sample(nrow(df)), x]  <- where x represents the number of rows that you want to select
```

## Exercise 10
### How could you put the columns in a data frame in alphabetical order?
```{r}
x <- c("b", "c", "a")
order(x) # changes to numeric
x[order(x)] # puts vector in alphabetical order

mtcars 
mtcars[order(names(mtcars))]

```
