---
title: "Wickham exercise"
author: "Qi Hao"
date: "2/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. what does dim() return when applied to a vector
```{r}
### it returns a vector
a <- rep(1,4)
dim(a) <- c(2,2)
class(a)
### Or an array
a <- rep(1,8)
dim(a) <- c(2,2,2)
class(a)
```

### 2. If is.matrix(x) is TRUE, what will is.array(x) return?
```{r}
### if is.matrix(x) is TRUE, is.array(x) would also return TRUE because matrix is a special kind of array
b <- matrix(1:4, nrow = 2, ncol = 2)
is.matrix(b)
is.array(b)
```

### 3. How would you describe the following three objects? What makes them different to 1:5?
```{r}
x1 <- array(1:5, c(1, 1, 5)) ### is a three dimension array, with 1:5 listed in the third dimension
x2 <- array(1:5, c(1, 5, 1)) ### is a three dimension array, with 1:5 listed in the second dimension
x3 <- array(1:5, c(5, 1, 1)) ### is a three dimension array, with 1:5 listed in the first dimension
### x1, x2, x3 differ from a vector 1:5 in that they have multiple dimensions although some of the diemsnions are contracted
```

### 1. What attributes does a data frame possess?
```{r}
### a data frame has attributes of colnames, rownames and class. for example:
dfm <- data.frame(x = 1:3, y = I(matrix(1:9, nrow = 3)))
attributes(dfm)
```

### 2. What does as.matrix() do when applied to a data frame with columns of different types?
```{r}
### it coercises the matrix into same type. For example, there is both integer and string, all is turned into string.
good <- data.frame(a = 1:2, b = c("a", "b"),
                   stringsAsFactors = FALSE)
bad <- as.matrix(good)
typeof(bad)
```

### 3. Can you have a data frame with 0 rows? What about 0 columns?
```{r}
### Yes you can. see below
zerodf <- data.frame(x = c(), y = c())
zerodf
```

### 1. Given a linear model, e.g., mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Extract the R squared from the model summary (summary(mod))
```{r}
mod <- lm(mpg ~ wt, data = mtcars)
str(summary(mod))
### extract residual degree of freedom
summary(mod)$df[2]
### extract the R squared
summary(mod)$r.squared
```
